\TESS Photometric Mapping of a Terrestrial Planet in the Habitable Zone: Detection of Clouds, Oceans, and Continents

### 2019-03-27 tex/ms.tex
better

We downloaded 
\footnote{\url{https://archive.stsci.edu/missions/tess/models/}}
the Solar System and \tess ephemeris data for
Sectors 1 and 2 and used \spiceypy \citep{Acton1996, Acton2017, Annex2017}
to convert the TJD timestamps to 
ephemeris time (ET). We then used the \textsf{spkezr} utility to compute
the relative positions of Sol, Sol d, and \tess for every cadence
in our dataset in the J2000 equatorial reference frame. In this right-handed
coordinate frame, Sol d is centered at the origin, the $y$ axis points along the 
planet's spin axis, and the $x$ axis points toward the vernal equinox. We
do not apply any light travel correction, as the expected delay is on the order
of a few seconds.

@@

Let us consider a right-handed coordinate frame in which the planet has unit 
radius and is located at the origin, the illumination source is along the $y-z$
plane, and the sub-stellar point is on the $+y$ axis at $x = 0$
(see Figure~\ref{fig:geometry}). In this frame, the terminator is a segment of
an ellipse, with semi-major axis $a=1$ aligned with the $x$ axis and
semi-minor axis $b$. It is straightforward to show that the sub-stellar point
is located at $y=\sqrt{1 - b^2}$ and the illumination is given by
%
\begin{align}
    I(b; x, y) &= 
    \begin{dcases}
        \sqrt{1 - b^2}y - bz & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            y \geq -b \sqrt{1 - x^2}
        \\
        0 & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            \mathrm{otherwise}
    \end{dcases}
    \quad ,
    \label{eq:illum}
\end{align}
%
where $z = \sqrt{1 - x^2 - y^2}$.
%
Since the illumination function is just a polynomial in $x$, $y$, and $z$,
weighting the terms in $\tilde{p}$ by this function keeps all terms within the
polynomial basis.
Provided we always rotate the problem such that the
sub-stellar point lies along the $+y$ axis as above, the limits of integration
are $-1 \leq x \leq 1$ and $b\sqrt{1 - x^2} \leq y \leq \sqrt{1 - x^2}$.
Computing the total reflected flux for any surface map is therefore a matter
of solving integrals of the form
%
\begin{align}
    J_{ijk}(b) &= \int_{-1}^{1} \int_{b\sqrt{1 - x^2}}^{\sqrt{1 - x^2}} x^i y^j z^k \mathrm{d} y \ \mathrm{d} x
    \nonumber\\
    &=
    \frac{
        \Gamma\left(\frac{i + 1}{2}\right) \Gamma\left(\frac{j + 1}{2}\right)
    }{
        2 \Gamma\left(\frac{i + j + k + 4}{2}\right)
    }
    \begin{dcases}
        0
        &
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        i \, \mathrm{odd}
        %
        \\
        %
            1 - b^{j + 1}
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 0
        %
        \\
        %
        \frac{\sqrt{\pi}}{2} - b^{j + 1}\Gamma\left(\frac{j + 4}{2}\right)
        {_2\tilde{F}_1} \left( -\frac{1}{2}, \frac{j + 1}{2}; \frac{j + 3}{2}; b^2\right)
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 1
    \end{dcases}
    \quad ,
    \label{eq:integral}
\end{align}
%
where ${_2\tilde{F}_1}$ is the regularized hypergeometric function. For all values
of $i$, $j$, and $k$ in $\tilde{p}$, ${_2\tilde{F}_1}$ reduces to simple
trigonometric functions involving $b$. Moreover, it is straightforward to
derive recursion relations for $J_{ijk}(b)$, making its evaluation extremely fast.
We implemented the algorithm for computing $J_{ijk}(b)$ in the development version
of \starry
\footnote{\url{https://github.com/rodluger/starry/tree/linear}}
and will be discussed in detail in upcoming work.

@@

As with the systematics model (\S\ref{sec:systematics}), the model
for the astrophysical signal is \emph{linear}: the signal $\mathbf{s}$ 
is a linear operation on the vector of spherical harmonic coefficients 
that describes the body's surface:
%
\begin{align}
    \mathbf{s} = \mathbf{A} \mathbf{y}
\end{align}
%
where $\mathbf{A}$ is the design matrix and $\mathbf{y}$ is the vector of spherical harmonic coefficients. 
Note that in \cite{Luger2019}, $\mathbf{y}$ describes the emissivity of the surface,
but here we will use it to represent the \emph{albedo} of the planet.
We choose a maximum spherical harmonics degree $l_{max} = 15$ for our fits,
noting that increasing $l_{max}$ above this value does not significantly
improve the quality of our fits. Note that $l = 15$ corresponds to
a maximum surface resolution on the order of $180^\circ/15 = 12^\circ$.

@@

Our full model for each light curve is simply the product of the 
systematics model $\mathbf{p}_n$ (different for each target) and
the \starry model $\mathbf{s}$ (shared by all targets). Since we happen
to know the exact distance $\mathbf{r}$ between \tess and Sol d at all times
(\S\ref{sec:orbit}), we also include the inverse square of $\mathbf{r}$ 
as a multiplicative term. The model for the flux timeseries
in the $n^\mathrm{th}$ postage stamp is therefore
%
\begin{align}
    \label{eq:model}
    \mathbf{m}_n = (\mathbf{r}^{-2}) \circ (\mathbf{B} \mathbf{w}_n) \circ (\mathbf{A} \mathbf{y})
\end{align}
%
where $\circ$ denotes the element-wise product of two vectors.

@@

In general, we find that the iterative scheme converges within about 50 
iterations, taking about three minutes on a laptop computer. However, since it 
is strongly regularized, the model significantly
underfits the data. In the second step, we remove our L2 prior on the spherical
harmonic coefficients and instead regularize the actual surface albedo, which
we compute on a uniform spherical grid with 50,000 points. We enforce a 
uniform prior in the range $[0, 1]$, with a small amount of Gaussian smoothing on either 
side. We additionally impose a constraint on the power spectrum of the
inferred map, requiring that it be drawn from a distribution whose mean
is a decaying power law in the spherical harmonic degree $l$ in order
to suppress ringing and spurious high order features.
Our full negative log likelihood function in this step is therefore
%
\begin{align}
    \label{eq:nll}
    -\log\mathcal{L} \, = \, 
        &\frac{1}{2}(\mathbf{f} - \mathbf{m})^\top \boldsymbol{\Sigma}^{-1} (\mathbf{f} - \mathbf{m}) \, + \nonumber \\
        &\frac{1}{2}\mathbf{w}^\top \boldsymbol{\Lambda}_w^{-1} \mathbf{w} \, + \nonumber \\
        &\frac{1}{2}(\mathbf{a_-} + 1 - \mathbf{a_+})^\top \boldsymbol{\Lambda}_a^{-1} (\mathbf{a_-} + 1 - \mathbf{a_+}) \, + \nonumber \\
        &\frac{1}{2}(\boldsymbol{\rho} - \boldsymbol{l}^{-\gamma})^\top \boldsymbol{\Lambda}_\rho^{-1} (\boldsymbol{\rho} - \boldsymbol{l}^{-\gamma}) \, + \nonumber \\
        &\frac{1}{2} \frac{(\gamma - \gamma_0)^2}{\sigma_\gamma^2}
        \quad,
\end{align}
%
where $\mathbf{f}$ is the measured flux across all targets,
$\mathbf{m}$ is the full model (Equation~\ref{eq:model}),
$\boldsymbol{\Sigma}$ is the (diagonal) data covariance, 
$\mathbf{w}$ are the systematics model weights for all targets,
$\boldsymbol{\Lambda}_w$ is the prior covariance of the weights
(Equation~\ref{eq:wprior}), $\mathbf{a_-}$ and $\mathbf{a_+}$ are
vectors containing the values of the albedo that are below zero and above one, 
respectively, $\boldsymbol{\Lambda}_a$ is the covariance of the Gaussian
used to smooth the edges of the top-hat prior on $\mathbf{a}$,
$\boldsymbol{\rho}$ is the power spectrum of the inferred spherical harmonic map
(equal to the sum of the squares of the coefficients at each degree $\boldsymbol{l}$),
$\gamma$ is the power law index of the power spectrum prior, $\boldsymbol{\Lambda}_\rho$ is the
covariance of that prior, $\gamma_0 = 1.5$ is the mean of the prior on $\gamma$
and $\sigma_\gamma^2 = 10^{-2}$ is the corresponding variance.
%
For simplicity, all our prior covariances are diagonal and
homoscedastic with variances equal to $10^{-2}$ for $\mathbf{w}$, and
$10^{-8}$ for $\mathbf{a}$, $10^{-1}$ for $\mathbf{\rho}$.

Unfortunately, the constraints outlined above break the linearity 
of the problem, and we must turn to a non-linear optimizer. We use the 
\textsf{AdamOptimizer} method in the \tf package \citep{Abadi2015}
to maximize Equation~(\ref{eq:nll}). Starting with the values of the
weights obtained in the first step, we run the \textsf{AdamOptimizer}
for 1000 iterations, well past convergence. Since \tf automatically
computes and propagates gradient information, this step is similarly fast
and also finishes in under three minutes on a standard computer.

@@

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/map.pdf}
    \caption{\label{fig:map}
             \emph{Top:} The maximum likelihood model for the surface albedo
             of Sol d, weighted by the visibility across Sectors 1 and 2. The
             dashed white lines indicate latitudes above/below which the
             visibility drops below 50\%. Black contours indicate coastlines
             derived from current state-of-the-art models of the surface geography 
             of Sol d.
             \emph{Bottom:} Actual corrected surface reflectance in the red band of the
             \emph{VIIRS} imager, averaged over the same date range, band limited
             to $l_\mathrm{max} = 15$, and weighted
             by the same visibility function. This bandpass covers only about one
             quarter of the \tess bandpass, so deviations between the two images
             are to be expected.
             \codelink{earthshine_S1_S2}
             }
    \end{centering}
\end{figure}

@@

The bottom panel of Figure~\ref{fig:map} shows the corrected surface
reflectance obtained from the
Visible Infrared Imaging Radiometer Suite (\emph{VIIRS}) instrument aboard the 
Suomi National Polar-orbiting Partnership (\emph{S-NPP}) satellite
\footnote{Data obtained via 
\href{https://github.com/rodluger/earthshine/blob/master/tex/figures/viirs.sh}{this wget script}.}. 
The image shows the corrected surface reflectance in the I1 band of the instrument,
averaged over the 28 days of data analyzed in this work, smoothed with a Gaussian
filter, and weighted by the \tess visibility as in the top panel.
The I1 band extends from 0.6 to 0.68 $\mu\mathrm{m}$
\footnote{\url{http://rammb.cira.colostate.edu/projects/npp/VIIRS_bands_and_bandwidths.pdf}},
and therefore covers only about one-fourth of the bandpass of \tess
(nominally 0.6 - 1.0 $\mu\mathrm{m}$).

@@

@@

>>>>

We downloaded 
\footnote{\url{https://archive.stsci.edu/missions/tess/models/}}
the Solar System and \tess ephemeris data for
Sectors 1 and 2 and used \spiceypy \citep{Acton1996, Acton2017, Annex2017}
to convert the TJD timestamps to 
ephemeris time (ET). We then used the \textsf{spkezr} utility of \spiceypy to compute
the relative positions of Sol, Sol d, and \tess for every cadence
in our dataset in the J2000 equatorial reference frame. In this right-handed
coordinate frame, Sol d is centered at the origin, the $y$ axis points along the 
planet's spin axis, and the $x$ axis points toward the vernal equinox. We
do not apply any light travel correction, as the expected delay is on the order
of a few seconds.

@@

Let us consider a right-handed coordinate frame in which the planet has unit 
radius and is located at the origin, the illumination source is along the $y-z$
plane, and the sub-stellar point is on the $+y$ axis at $x = 0$
(see Figure~\ref{fig:geometry}). In this frame, the terminator is a segment of
an ellipse, with semi-major axis $a=1$ aligned with the $x$ axis and
semi-minor axis $b$. It is straightforward to show that the sub-stellar point
is located at $y=\sqrt{1 - b^2}$ and the illumination is given by
%
\begin{align}
    I(b; x, y) &= 
    \begin{dcases}
        \sqrt{1 - b^2}y - bz & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            y \geq -b \sqrt{1 - x^2}
        \\
        0 & 
            \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
            \mathrm{otherwise}
    \end{dcases}
    \quad ,
    \label{eq:illum}
\end{align}
%
where $z = \sqrt{1 - x^2 - y^2}$.
%
Since the illumination function is just a polynomial in $x$, $y$, and $z$,
weighting the terms in $\tilde{p}$ by this function keeps all terms within the
polynomial basis.
Provided we always rotate the problem such that the
sub-stellar point lies along the $+y$ axis as above, the limits of integration
are $-1 \leq x \leq 1$ and $b\sqrt{1 - x^2} \leq y \leq \sqrt{1 - x^2}$.
Computing the total reflected flux for any surface map is therefore a matter
of solving integrals of the form
%
\begin{align}
    J_{ijk}(b) &= \int_{-1}^{1} \int_{b\sqrt{1 - x^2}}^{\sqrt{1 - x^2}} x^i y^j z^k \mathrm{d} y \ \mathrm{d} x
    \nonumber\\
    &=
    \frac{
        \Gamma\left(\frac{i + 1}{2}\right) \Gamma\left(\frac{j + 1}{2}\right)
    }{
        2 \Gamma\left(\frac{i + j + k + 4}{2}\right)
    }
    \begin{dcases}
        0
        &
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        i \, \mathrm{odd}
        %
        \\
        %
            1 - b^{j + 1}
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 0
        %
        \\
        %
        \frac{\sqrt{\pi}}{2} - b^{j + 1}\Gamma\left(\frac{j + 4}{2}\right)
        {_2\tilde{F}_1} \left( -\frac{1}{2}, \frac{j + 1}{2}; \frac{j + 3}{2}; b^2\right)
        & 
        \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad 
        k = 1
    \end{dcases}
    \quad ,
    \label{eq:integral}
\end{align}
%
where ${_2\tilde{F}_1}$ is the regularized hypergeometric function. For all values
of $i$, $j$, and $k$ in $\tilde{p}$, ${_2\tilde{F}_1}$ reduces to simple
trigonometric functions involving $b$. Moreover, it is straightforward to
derive recursion relations for $J_{ijk}(b)$, making its evaluation extremely fast.
We implemented the algorithm for computing $J_{ijk}(b)$ in the development version
of \starry
\footnote{\url{https://github.com/rodluger/starry/tree/linear}};
the details will be discussed in more detail in upcoming work.

@@

As with the systematics model (\S\ref{sec:systematics}), the model
for the astrophysical signal is \emph{linear}: the signal $\mathbf{s}$ 
is a linear operation on the vector of spherical harmonic coefficients 
$\mathbf{y}$
that describes the body's surface:
%
\begin{align}
    \mathbf{s} = \mathbf{A} \mathbf{y}
\end{align}
%
where $\mathbf{A}$ is the design matrix and $\mathbf{y}$ is the vector of spherical harmonic coefficients. 
Note that in \cite{Luger2019}, $\mathbf{y}$ describes the emissivity of the surface,
but here we will use it to represent the \emph{albedo} of the planet.
We choose a maximum spherical harmonic degree $l_{max} = 25$ for our fits,
noting that increasing $l_{max}$ above this value does not significantly
improve the quality of our fits. Note that $l = 25$ corresponds to
a maximum surface resolution on the order of $180^\circ/25 \approx 7^\circ$.

@@

Our full model for each light curve is simply the product of the 
systematics model $\mathbf{p}_n = \mathbf{B}\mathbf{w}_n$ (different for each target) and
the \starry model $\mathbf{s} = \mathbf{A}\mathbf{y}$ (shared by all targets). Since we happen
to know the exact distance $\mathbf{r}$ between \tess and Sol d at all times
(\S\ref{sec:orbit}), we also include the inverse square of $\mathbf{r}$ 
as a multiplicative term. The model for the flux timeseries
in the $n^\mathrm{th}$ postage stamp is therefore
%
\begin{align}
    \label{eq:model}
    \mathbf{m}_n = (\mathbf{r}^{-2}) \circ (\mathbf{B} \mathbf{w}_n) \circ (\mathbf{A} \mathbf{y})
\end{align}
%
where $\circ$ denotes the element-wise product of two vectors.

@@

In general, we find that the iterative scheme converges within about 50 
iterations, taking about fifteen minutes on a laptop computer. However, since it 
is strongly regularized, the model significantly
underfits the data. In the second step, we remove our L2 prior on the spherical
harmonic coefficients and instead regularize the actual surface albedo, which
we compute on a uniform spherical grid with 50,000 points. We enforce a 
uniform prior in the range $[0, 1]$, with a small amount of Gaussian smoothing on either 
side. We additionally impose a constraint on the power spectrum of the
inferred map, requiring that it be drawn from a distribution whose mean
is a decaying power law in the spherical harmonic degree $l$ in order
to suppress ringing and spurious high order features.
Our full negative log likelihood function in this step is therefore
%
\begin{align}
    \label{eq:nll}
    -\log\mathcal{L} \, = \, 
        &\frac{1}{2}(\mathbf{f} - \mathbf{m})^\top \boldsymbol{\Sigma}^{-1} (\mathbf{f} - \mathbf{m}) \, + \nonumber \\
        &\frac{1}{2}\mathbf{w}^\top \boldsymbol{\Lambda}_w^{-1} \mathbf{w} \, + \nonumber \\
        &\frac{1}{2}(\mathbf{a_-} + 1 - \mathbf{a_+})^\top \boldsymbol{\Lambda}_a^{-1} (\mathbf{a_-} + 1 - \mathbf{a_+}) \, + \nonumber \\
        &\frac{1}{2}(\boldsymbol{\rho} - \boldsymbol{l}^{-\gamma})^\top \boldsymbol{\Lambda}_\rho^{-1} (\boldsymbol{\rho} - \boldsymbol{l}^{-\gamma}) \, + \nonumber \\
        &\frac{1}{2} \frac{(\gamma - \gamma_0)^2}{\sigma_\gamma^2}
        \quad,
\end{align}
%
where $\mathbf{f}$ is the measured flux across all targets,
$\mathbf{m}$ is the full model (Equation~\ref{eq:model}),
$\boldsymbol{\Sigma}$ is the (diagonal) data covariance, 
$\mathbf{w}$ are the systematics model weights for all targets,
$\boldsymbol{\Lambda}_w$ is the prior covariance of the weights
(Equation~\ref{eq:wprior}), $\mathbf{a_-}$ and $\mathbf{a_+}$ are
vectors containing the values of the albedo that are below zero and above one, 
respectively, $\boldsymbol{\Lambda}_a$ is the covariance of the Gaussian
used to smooth the edges of the top-hat prior on $\mathbf{a}$,
$\boldsymbol{\rho}$ is the power spectrum of the inferred spherical harmonic map
(equal to the sum of the squares of the coefficients at each degree $\boldsymbol{l}$),
$\gamma$ is the power law index of the power spectrum prior, $\boldsymbol{\Lambda}_\rho$ is the
covariance of that prior, $\gamma_0 = 1.5$ is the mean of the prior on $\gamma$
and $\sigma_\gamma^2 = 10^{-2}$ is the corresponding variance.
%
For simplicity, all our prior covariances are diagonal and
homoscedastic with variances equal to $10^{-2}$ for $\mathbf{w}$, and
$10^{-8}$ for $\mathbf{a}$, $10^{-1}$ for $\boldsymbol{\rho}$.

Unfortunately, the constraints outlined above break the linearity 
of the problem, and we must turn to a non-linear optimizer. We use the 
\textsf{AdamOptimizer} method in the \tf package \citep{Abadi2015}
to maximize Equation~(\ref{eq:nll}). Starting with the values of the
weights obtained in the first step, we run the \textsf{AdamOptimizer}
for 1000 iterations, well past convergence. Since \tf automatically
computes and propagates gradient information, this step is fast
and also finishes in under five minutes on a standard computer.

@@

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/map.pdf}
    \caption{\label{fig:map}
             \emph{Top:} The maximum likelihood model for the surface albedo
             of Sol d, weighted by the visibility across Sectors 1 and 2. The
             dashed white lines indicate latitudes above/below which the
             visibility drops below 50\%. Black contours indicate coastlines
             derived from current state-of-the-art models of the surface geography 
             of Sol d.
             \emph{Bottom:} Approximate cloud coverage map based on corrected
             surface reflectance obtained by the
             \emph{VIIRS} imager, averaged over the same date range and weighted
             by the same visibility function.
             \codelink{earthshine_S1_S2}
             }
    \end{centering}
\end{figure}

@@

The bottom panel of Figure~\ref{fig:map} shows an approximate average
cloud cover map of Sol d based on imagery taken by the
Visible Infrared Imaging Radiometer Suite (\emph{VIIRS}) instrument aboard the 
Suomi National Polar-orbiting Partnership (\emph{S-NPP}) satellite
\footnote{Data obtained via 
\href{https://github.com/rodluger/earthshine/blob/master/tex/figures/viirs.sh}{this wget script}.}. 
The image was produced by \todo{explain...}
averaged over the 28 days of data analyzed in this work, smoothed with a Gaussian
filter, and weighted by the \tess visibility as in the top panel.

@@

\subsection{Temporal Evolution}

@@

\subsection{Sources of Bias}

In our analysis, we neglected the effects of limb darkening. Because of
atmospheric attenuation, features near the limb of the planet should in
reality contribute less to the flux than our model predicts. This could
bias our results particularly in cases where the planet is seen at half
phase (for instance, during days 1326 and 1341 in Figure~\ref{fig:phases}).
In the absence of limb darkening, the brightest regions on the planet are those
close to the sub-stellar point, which is on the limb. Limb darkening should
shift the regions of peak brightness closer to the center of the planet, adding
an east-west bias in the location of the features we identify.

We also ignored the finite angular size of Sol at Sol d (about $0.5^\circ$),
treating it instead as a point source. Regions immediately nightward of the 
terminator should therefore contribute a small amount of flux, which we neglect.
The effect, however, is below the resolution of our inferred map, so we expect
it does not introduce significant bias.

Finally, we implicitly assumed an isotropic prior on the spherical harmonic
coefficients by imposing a power spectrum in only the degree $l$. This has the
effect of lessening the extent to which features can be confined to certain
latitudes. In particular, our prior implicitly disfavors banded cloud structure,
which one may \emph{a priori} expect on a rapidly rotating planet such as 
Sol d. The bottom panel of Figure~\ref{fig:map} shows such banded structure
at the equator and at mid-latitudes due to the combined effects of 
north-south circulation and deflection by the Coriolis force.
